---
title: "Senescence project"
author: "Amparo Daza"
date: "2025-02-13"
output: html_document
---
#####
```{r include=FALSE}

# Senescence analusis
## Libraries
library(dplyr)
library(lme4)
library(ggplot2)
library(tidyr)

## Set up directory
setwd("/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project")

# Load files 
data_10_11_23  <- read.csv("/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results/results_10_11_23_v1.csv")
```

```{r, echo = FALSE}

# In this code the structure was checked, change columns to factors for the model and group by Genotype and Repetition
# Check structure 
str(data_10_11_23)

# change columns to factor
data_10_11_23$Genotype <- as.factor(data_10_11_23$Genotype)
data_10_11_23$Rep <- as.factor(data_10_11_23$Rep)

data_10_11_23_plotbase_all <- data_10_11_23 %>%
  group_by(Genotype,Rep) %>%
  summarise(avg_Red = mean(Red),
            avg_Green = mean(Green),
            avg_Blue = mean(Blue),
            avg_BI = mean(BI),
            avg_GLI = mean(GLI),
            avg_NGRDI = mean(NGRDI),
            avg_VARI=mean(VARI),
            avg_BGI = mean(BGI),
            avg_BI_2 = mean(BI_2),
            avg_GLI_2 = mean(GLI_2),
            avg_NGRDI_2 = mean(NGRDI_2),
            avg_VARI_2=mean(VARI_2),
            avg_BGI_2 = mean(BGI_2),
            avg_BCC=mean(BCC),
            avg_CIVE=mean(CIVE)
            )

```



#### Fit the model for Plant Height 
```{r}
fit.CIVE <- lmer(avg_CIVE ~ (1|Genotype) + (1|Rep), data = data_10_11_23_plotbase_all) # Fit model for Plantheigh for exp1
summary(fit.CIVE)

```

#### Calculate h2 for Plant Height

```{r}

# Calculate heritability for Plant Height, the values were extracted from the Model

heritability_X = (2.146e-05/(2.146e-05+(9.373e-06/2)))
```


```{r, echo = FALSE}
# Calculate heritability for just the two inner rows 
# Heritability just 2 inner rows
data_10_11_23_plotbase_internal <- data_10_11_23 %>%
  filter(!(substr(PlotID_rowNumber, nchar(PlotID_rowNumber), nchar(PlotID_rowNumber)) %in% c("1","4")))

str(data_10_11_23_plotbase_internal)

#Heritabilities just internal

data_10_11_23_internal_avg <- data_10_11_23_plotbase_internal %>%
  group_by(Genotype,Rep) %>%
  summarise(avg_Red = mean(Red),
            avg_Green = mean(Green),
            avg_Blue = mean(Blue),
            avg_BI = mean(BI),
            avg_GLI = mean(GLI),
            avg_NGRDI = mean(NGRDI),
            avg_VARI=mean(VARI),
            avg_BGI = mean(BGI),
            avg_BI_2 = mean(BI_2),
            avg_GLI_2 = mean(GLI_2),
            avg_NGRDI_2 = mean(NGRDI_2),
            avg_VARI_2=mean(VARI_2),
            avg_BGI_2 = mean(BGI_2),
            avg_BCC=mean(BCC),
            avg_CIVE=mean(CIVE)
  )
```

#### Fit a new model with Genotype and Rep as random and calculate heritabilities for just the two inner rows for exp1
```{r}
fit.CIVE <- lmer(avg_CIVE ~ (1|Genotype) + (1|Rep), data = data_10_11_23_internal_avg) # Fit model for Plantheigh for exp1
summary(fit.CIVE)

#Calculate heritability for Plant Height
heritability_X= (2.508e-05/(2.508e-05+(1.044e-05/2)))
```

#### Plot heritabilities for the results for RGB data (All vegetative indexes)
```{r, echo = FALSE}

######### Plots

# 1. Plot 1 Heritabilities
## Load heritabilities 4 row plot:

### set the directory for the RGB data
setwd("/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results")

# Load files 
Plot1_h2_full<-read.csv("heritabilities_all.csv")

#Remove column 2
GLI <- Plot1_h2_full[,3]
NGRDI <- Plot1_h2_full[,4]
VARI <- Plot1_h2_full[,5]
BCC <- Plot1_h2_full[,6]
VICE <- Plot1_h2_full[,7]
dates <- as.Date(c("2023-8-10", "2023-8-16", "2023-8-23", "2023-8-30", "2023-9-05",
                   "2023-9-13", "2023-9-20", "2023-9-29", "2023-10-4", "2023-10-11"))

# Create a data frame
Plot1_h2 <- data.frame(dates,GLI, NGRDI, VARI, BCC,VICE)

# Convert data to long format for easier plotting

df_long <- gather(Plot1_h2, key = "VegetativeIndex", value = "Value", -dates)

# Plotting using ggplot2
ggplot(data = df_long, aes(x = dates, y = Value, color = VegetativeIndex)) +
  geom_line() +
  geom_point() +
  labs(x = "Date", y = "Value", title = "Vegetative Index over Time") +
  theme_minimal() +
  scale_color_discrete(name = "Vegetative Index")+
  labs(x = "Date of flight", y = "Vegetative Index heritability", title = "Heritability over Time")+
  theme(axis.text = element_text(size = 12))+
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

#### Comparison of heritabilities for each of the vegetative indices for the 4 rows vs the two inner rows
##### NGDRI Index

```{r, echo = FALSE}
# 2. Comparisson of h2
# NGRDI
## Sample data (replace with your actual data)
time <- as.Date(c("2023-8-10", "2023-8-16", "2023-8-23", "2023-8-30", "2023-9-05",
                  "2023-9-13", "2023-9-20", "2023-9-29", "2023-10-4", "2023-10-11"))
method <- rep(c("4 Rows", "2 Inner rows"), each = 10)
value <- c(0.8,0.84,0.91,0.82,0.9,0.88,0.86,0.83,0.6
           ,0.41,0.75,0.83,0.91,0.81,0.9,0.89,0.86,0.82,0.69,0.32)
             # Replace with your actual data

# Create a data frame
df <- data.frame(time, method, value)


ggplot(df, aes(x = time, y = value, fill = method)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  labs(x = "Date of flight", y = "Vegetative Index heritability", title = "NGRDI", 
       subtitle = "4 vs 2 Inner rows") +
  scale_fill_manual(values = c("#FF9999", "#66B2FF"), name = "Method") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_blank(), axis.line = element_line(colour = "black"),
        text = element_text(size = 12, family = "Arial"),
        plot.title = element_text(size = 20, hjust = 0.5, face = "bold.italic"),
        plot.subtitle = element_text(size = 16, hjust = 0.5, face = "italic"),
        plot.caption = element_text(size = 10, hjust = 1),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.margin = unit(c(1, 4, 1, 1), "cm"))
```

##### BCC Index

```{r, echo = FALSE}
# BCC
## Sample data (replace with your actual data)
time <- as.Date(c("2023-8-10", "2023-8-16", "2023-8-23", "2023-8-30", "2023-9-05",
                  "2023-9-13", "2023-9-20", "2023-9-29", "2023-10-4", "2023-10-11"))
method <- rep(c("4 Rows", "2 Inner rows"), each = 10)
value <- c(0.7,0.79,0.87,0.83,0.65,0.78,0.72,0.9,
           0.84,0.49,0.67,0.77,0.83,0.8,0.6,0.76,0.7,0.87,0.79,0.63)
# Replace with your actual data

# Create a data frame
df <- data.frame(time, method, value)


ggplot(df, aes(x = time, y = value, fill = method)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  labs(x = "Date of flight", y = "Vegetative Index heritability", title = "BCC", 
       subtitle = "4 vs 2 Inner rows") +
  scale_fill_manual(values = c("#FF9999", "#66B2FF"), name = "Method") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_blank(), axis.line = element_line(colour = "black"),
        text = element_text(size = 12, family = "Arial"),
        plot.title = element_text(size = 20, hjust = 0.5, face = "bold.italic"),
        plot.subtitle = element_text(size = 16, hjust = 0.5, face = "italic"),
        plot.caption = element_text(size = 10, hjust = 1),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.margin = unit(c(1, 4, 1, 1), "cm"))
```

##### CIVE Index
```{r, echo = FALSE}
#CIVE

time <- as.Date(c("2023-8-10", "2023-8-16", "2023-8-23", "2023-8-30", "2023-9-05",
                  "2023-9-13", "2023-9-20", "2023-9-29", "2023-10-4", "2023-10-11"))
method <- rep(c("4 Rows", "2 Inner rows"), each = 10)
value <- c(0.84,0.86,0.87,0.83,0.71,0.76,0.73,0.82,0.84,0.82,
           0.8,0.85,0.83,0.8,0.69,0.77,0.7,0.81,0.81,0.83)
# Replace with your actual data

# Create a data frame
df <- data.frame(time, method, value)


ggplot(df, aes(x = time, y = value, fill = method)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  labs(x = "Date of flight", y = "Vegetative Index heritability", title = "CIVE", 
       subtitle = "4 vs 2 Inner rows") +
  scale_fill_manual(values = c("#FF9999", "#66B2FF"), name = "Method") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_blank(), axis.line = element_line(colour = "black"),
        text = element_text(size = 12, family = "Arial"),
        plot.title = element_text(size = 20, hjust = 0.5, face = "bold.italic"),
        plot.subtitle = element_text(size = 16, hjust = 0.5, face = "italic"),
        plot.caption = element_text(size = 10, hjust = 1),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_text(size = 10)) +
  theme(plot.margin = unit(c(1, 4, 1, 1), "cm"))
```

##### Analysis of the Stay green vs VI

```{r, echo = FALSE}

################## Stay green vs VI

setwd("/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results")

# Load files 
Staygreen <- read.csv("staygreen_by_Index.csv")
Staygreen[8,4] <- NA
str(Staygreen)
Staygreen$NGRDI <- as.numeric(Staygreen$NGRDI)
```

```{r, echo = FALSE}
# Create scatterplot with trendline and R-squared
model <- lm(NGRDI ~ Green_PCT, data = Staygreen)
rsquared <- summary(model)$r.squared

ggplot(Staygreen, aes(x = BCC, y = Green_PCT)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Vegetation Index Value", y = "Staygreen score", 
       title = "BCC vs Staygreen",
       subtitle = paste("r = ","0.43")) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r, echo = FALSE}
###########
# NGRDI curves

# Load files 
geno_summary2 <- read.csv("/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results/NGRVI_geno_summary2.csv")

data_long <- gather(geno_summary2, NGRDV, value, X90:X152, factor_key = TRUE)

#data_long$NGRDV <- substr(data_long$NGRDV, 2, nchar(data_long$NGRDV))

data_long$NGRDV <- as.numeric(data_long$NGRDV)
data_long$NGRDV[data_long$NGRDV == "1"]<- "90"
data_long$NGRDV[data_long$NGRDV == "2"]<- "96"
data_long$NGRDV[data_long$NGRDV == "3"]<- "103"
data_long$NGRDV[data_long$NGRDV == "4"]<- "110"
data_long$NGRDV[data_long$NGRDV == "5"]<- "116"
data_long$NGRDV[data_long$NGRDV == "6"]<- "124"
data_long$NGRDV[data_long$NGRDV == "7"]<- "131"
data_long$NGRDV[data_long$NGRDV == "8"]<- "140"
data_long$NGRDV[data_long$NGRDV == "9"]<- "152"

data_long$NGRDV <- as.numeric(data_long$NGRDV)

ggplot(data=data_long, aes(x=NGRDV, y=value, color=Genotype)) +
   geom_line()+
  scale_color_brewer(palette="Paired")+
  theme_minimal() +
  theme(legend.position = "none")
```


```{r, echo = FALSE}
#Data long 2

data_long2 <- data_long
df_long2 <- subset(data_long, grepl("^D", Genotype))

ggplot(data=df_long2, aes(x=NGRDV, y=value, color=Genotype)) +
  geom_line()+
  theme_minimal()+
  labs(x = "Days after planting", y = "NGRDI Index Value", title = "NGRDI post anthesis for Checks")+
  theme(axis.text = element_text(size = 12))+
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

```{r include = FALSE}
  scale_fill_identity()



#Add stay green
  df_long2$staygreen = (df_long2$value - 0.0272797)/0.0005933
```


```{r, echo = FALSE}
#-------------
  # Black Layer
  
BL_Master <- read.csv('/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results/BL_Master.csv')
```


##### Fit model with RM for each Genotype and days form planting to Black layer

```{r}
  # Create scatterplot with trendline and R-squared
model <- lm(Days_Planting_to_BL ~ Average.of.RM, data = BL_Master)
rsquared <- summary(model)$r.squared
  
```

```{r, echo = FALSE}
ggplot(BL_Master, aes(x = Days_Planting_to_BL, y = Average.of.RM)) +
    geom_point(position = position_jitter(width = 0.15, height = 0.15), size = 1) +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(x = "Days from Planting to Black Layer", y = "RM", 
         title = "Days to Blacklayer vs RM",
         subtitle = paste("r = ","0.46")) +
    theme_minimal()+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.title = element_text(size = 16),
          axis.text = element_text(size = 16))

subset_commercial <- subset(BL_Master, 
                            Commercial == "YES")
```

##### Fit the model Average Relative Maturity for each comercial genotype and Days from planting to Black Layer

```{r}
model <- lm(Days_Planting_to_BL ~ Average.of.RM, data = subset_commercial)
rsquared <- summary(model)$r.squared

```

```{r, echo = FALSE}


ggplot(subset_commercial, aes(x = Days_Planting_to_BL, y = Average.of.RM, color=Genotype)) +
  geom_point(position = position_jitter(width = 0.15, height = 0.15), size = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Days from Planting to Black Layer", y = "RM", 
       title = "Days to Blacklayer vs RM",
       subtitle = paste("r = ","0.8")) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 16))
```

#---------------

##### Fit the model Average Relative Maturity for each comercial genotype and Days from planting to 50% pollen

```{r}
model <- lm(Days_Planting_to_Pollen ~ Average.of.RM, data = subset_commercial)
rsquared <- summary(model)$r.squared
```

```{r, echo = FALSE}
ggplot(subset_commercial, aes(x = Days_Planting_to_Pollen, y = Average.of.RM, color=Genotype)) +
  geom_point(position = position_jitter(width = 0.15, height = 0.15), size = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Days from Planting Pollen", y = "RM", 
       title = "Days from planting to Pollen vs RM",
       subtitle = paste("r = ","0.95")) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 16))
```


##### Fit model Average Relative Maturity for each genotype and Days from planting to Pollen

```{r}
model <- lm(Average.of.RM ~ Days_Planting_to_Pollen, data = BL_Master)
rsquared <- summary(model)$r.squared
```

```{r, echo = FALSE}
ggplot(BL_Master, aes(x = Days_Planting_to_Pollen, y = Average.of.RM)) +
  geom_point(position = position_jitter(width = 0.15, height = 0.15), size = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Days Planting to pollen", y = "RM", 
       title = "Days from planting to Pollen to vs RM",
       subtitle = paste("r = ","0.6")) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 16))
```


##### Moisture
```{r, echo = FALSE}
MST <- read.csv('/Users/amparodaza/Library/CloudStorage/OneDrive-UW-Madison/Doctorado/Senescence project/RGB_Data/All_results/Final.csv')
```

##### Fit the model with CIVE Index and Moisture
```{r}
## CIVE use to get plant biomass and distinguish plants from soil and residue backgrounds
# Create scatterplot with trendline and R-squared
model <- lm(Moisture ~ CIVE, data = MST)
rsquared <- summary(model)$r.squared
```

```{r, echo = FALSE}
ggplot(MST, aes(x = Moisture, y = CIVE)) +
  geom_point(position = position_jitter(width = 0.15, height = 0.15), size = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Kernel Moisture", y = "CIVE Value", 
       title = "CIVE vs Kernel Moisture %",
       subtitle = paste("r = ","0.43")) +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 16))

```

######-----------------

##### Yield calculation and Heatmap

$Adj \space weight \space (lb) = \frac{(Weight \space (lb)-(Weight \space (lb)*Moisture)}{0.845}$

$Yield \space (bu/ac) = Adj \space Weight \space (lb) *\frac{1 \space bushel}{1 \space acre}*\frac{43560 \space ft^2}{1 \space acre}*\frac{1}{22.5 \space ft * 2.5 \space ft *2}$



```{r, echo = FALSE}

BL_Master <- BL_Master %>%
  group_by(Genotype) %>%
  mutate(GenotypeNumber = cur_group_id()) %>%  # Creates a unique ID for each Genotype
  ungroup()

midpoint <- 11

ggplot(BL_Master, aes(x = Row, y = Range, fill = Yield)) + 
  geom_tile(color = "black") + 
  geom_text(aes(label = paste(GenotypeNumber, "=", round(Yield))), color = "white", size = 3) +
  scale_fill_gradient(low = "red", high = "green") +
  labs(x = "Row", y = "Range", fill = "Yield") +
  geom_hline(yintercept = midpoint + 0.5, color = "black", linewidth = 1) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(BL_Master$Row), max(BL_Master$Row), by = 1)) +
  theme(axis.text.x = element_text(hjust = 1),  # Rotate x-axis labels
        panel.grid.major = element_blank(), # Remove grid lines
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank())
```
